---
title: 'CMTH 642 Data Analytics: Advanced Methods'
output:
  pdf_document: default
  html_document: default
  word_document: default
---
<center> <h1> Assignment 2 (10%) </h1> </center>
<center>  <h3> [Ilhak Park] </h2> </center>
<center> <h3> [DJ0 & 501072432] </h2> </center>
---

#### 1. Read the csv file (USDA_Clean.csv) in the folder and assign it to a data frame. (3 points)
```{r}
USDAClean <-read.csv(file="USDA_Clean.csv", sep =",")
```

#### 2. Check the datatypes of the attributes. (3 points)
```{r}
sapply (USDAClean, class)
```

#### 3. Visualize the correlation among Calories, Protein, Total Fat, Carbohydrate, Sodium and Cholesterol. (7 points)
```{r}
subset <- USDAClean[,c("Calories","Protein","TotalFat","Carbohydrate", "Sodium","Cholesterol")]
cor(subset)
library(corrplot)
corrplot(cor(subset), method="circle")
```

#### 4. Is the correlation between Calories and Total Fat statistically significant? Why? (7 points)
```{r}
cor (USDAClean$Calories, USDAClean$TotalFat)
#The correlation between Calories and Total Fat statistically significant because
#their correlation coefficient is about 0.804495 which is close to a perfect positive correlation, 1.0. 
```

#### 5. Create a Linear Regression Model, using Calories as the dependent variable Protein, Total Fat, Carbohydrate, Sodium and Cholesterol as the independent variables. (7 points)
```{r}
Calories_lm <- lm(Calories~ Protein+TotalFat+Carbohydrate+Sodium+Cholesterol, data= USDAClean)
Calories_lm
```

#### 6. Write the Linear Regression Equation, using Calories as the dependent variable whereas Protein, TotalFat, Carbohydrate, Sodium and Cholesterol as the independent variables. (7 points)
```{r}
Calories_lm$coefficients
#Calories=3.9882752613+3.9891994394*Protein+8.7716980068*TotalFat
#+3.7432000604*Carbohydrate+0.0003383021*Sodium+0.0110138110*Cholesterol
```

#### 7. Which independent variable is the least significant? Why? (7 points)
```{r}
anova(Calories_lm)
#The sodium is the least significant because it has a p-value of 0.1378, 
#while independent variables have much smaller p-values. 
#When an independent variable has a p-value of 0.1378, 
#it is difficult to predict changes to the data. 
```

#### 8. A new product is just produced with the following data:                                                            Protein=0.1, TotalFat=35, Carbohydrate=405, Sodium=440, Cholesterol=70, Sugar=NA, Calcium=35, Iron=NA, Potassium=35, VitaminC=10,	VitaminE=NA, VitaminD=NA.                                                                                                         Based on the model you created, what is the predicted value for Calories? (7 points)
```{r}
NewData = data.frame(Protein=0.1, TotalFat=35, Carbohydrate=405, Sodium=440, Cholesterol=70, Calcium=35,  Potassium=35, VitaminC=10)
predict(Calories_lm, NewData)
#The predicted value for Calories is 1828.312.
```

#### 9. If the Sodium amount increases from 440 to 44440 (10000% increase), how much change will occur on Calories in percent? Explain why? (7 points)
```{r}
NewData2 = data.frame(Protein=0.1, TotalFat=35, Carbohydrate=405, Sodium=44440, Cholesterol=70, Calcium=35,  Potassium=35, VitaminC=10)
predict(Calories_lm, NewData2)
(1843.198-1828.312)/1828.12 *100
#The Calories will increase by 0.8142792%. Since Sodium is the least significant independent variable, 
#the Calories has only increased by a small amount even though the Sodium amount has increased by 10000%.
```


#### 10. A study of primary education asked elementaty school students to retell two book articles that they read earlier in the week. The first (Article 1) had no pictures, and the second (Article 2) was illustrated with pictures. An expert listened to recordings of the students retelling each article and assigned a score for certain uses of language. Higher scores are better. Here are the data for five readers in this study:

#### Article 1   0.40   0.72   0.00   0.36   0.55

#### Article 2   0.77   0.49   0.66   0.28   0.38


#### A) What are $H_0$ and $H_a$ ? (5 points)
```{r}
#H_0: The population means of the scores from the two groups are equal, mu_1 = mu_2. 
#The null hypothesis states that illustrations do not affect how students retell the article.
#H_a: The population means of the scores from the two groups are not equal, mu_1 != mu_2. 
#The alternative hypothesis states that illustrations affect how students retell the article.
```

#### B) Is this a paired or unpaired experiment? (5 points)
```{r}
#This is a paired experiment because both article 1 and 2 are tested for each student.
```

#### C) Based on your previous answer, which nonparametric test statistic would you use to compare the medians of Article 1 and Article 2. (5 points) 
```{r}
#Since comparing two groups with nonparametric statistics and paired experiment, 
#Wilcoxon signed rank test should be used. 
```

#### D) Use a nonparametric test statistic to check if there is a statistically significant difference between the medians of Article 1 and Article 2. (5 points) 
```{r}
article1 <- c(0.40,0.72,0.00,0.36,0.55)
article2 <- c(0.77,0.49,0.66,0.28,0.38)
wilcox.test(article1, article2, paired=T)
```

#### E) Will you accept or reject your Null Hypothesis? ($\alpha = 0.05$) Do illustrations improve how the students retell an article or not? Why? (5 points)
```{r}
#Since p-value is greater than 0.05, we do not reject the null hypothesis; 
#therefore, illustrations do not improve how the students retell an article. 
#We accept the null hypothesis that the population means of the scores from the two groups are equal 
#because given the confidence level is 0.05, the p-value, 0.8125, is much greater.
```

#### 11. Two companies selling toothpastes with the lable of 100 grams per tube on the package. We randomly bought eight toothpastes from each company A and B from random stores. Afterwards, we scaled them using high precision scale. Our measurements are recorded as follows:

#### Company A: 97.1   101.3   107.8   101.9   97.4    104.5   99.5    95.1
#### Company B: 103.5  105.3   106.5   107.9   102.1   105.6   109.8   97.2


#### A) Is this a paired or unpaired experiment? (5 points)
```{r}
#This is an unpaired experiment because the toothpastes A and B 
#are not bought from the same store. 
```

#### B) Based on your previous answer, which nonparametric test statistic would you use to compare the medians of Company A and Company B. (5 points) 
```{r}
#Since comparing two groups with nonparametric statistics and unpaired experiment, 
#Wilcoxon rank sum test should be used.
```

#### C) Use a nonparametric test statistic to check if there is a statistically significant difference between the medians of Company A and Company B. (5 points) 
```{r}
CompanyA <- c(97.1,   101.3,   107.8,   101.9,   97.4,    104.5,   99.5,    95.1)
CompanyB <- c(103.5,  105.3,   106.5,   107.9,   102.1,   105.6,   109.8,   97.2)
wilcox.test(CompanyA, CompanyB, paired = FALSE)
```

#### D) Will you accept or reject your Null Hypothesis? ($\alpha = 0.05$) Are packaging process similar or different based on weight measurements? Why? (5 points)
```{r}
#Since the p-value is less than 0.05, we reject the Null Hypothesis.
#The packaging process between the two companies are different based on weight measurement 
#because we reject the Null Hypothesis that the two population means are the same. 
```

This is the end of Assignment 2

Ceni Babaoglu, PhD
