---
title: 'CMTH 642 Data Analytics: Advanced Methods'
output:
  pdf_document: default
  word_document: default
  html_document: default
---
<center> <h1> Assignment 1 (10%) </h1> </center>
<center>  <h3> [Ilhak Park] </h2> </center>
<center> <h3> [DJ0 & 501072432] </h2> </center>
---

#### 1. Read the csv files in the folder. (3 points)
```{r}
micro <- read.csv(file= "USDA_Micronutrients.csv", sep= ",")
macro <- read.csv(file="USDA_Macronutrients.csv", sep =",")
```

#### 2. Merge the data frames using the variable "ID". Name the Merged Data Frame "USDA". (6 points)
```{r}
USDA <- merge(macro,micro)
summary(USDA)
```

#### 3. Check the datatypes of the attributes. Delete the commas in the Sodium and Potasium records. Assign Sodium and Potasium as numeric data types. (6 points)
```{r}
sapply (USDA, class)
USDA$Sodium <- gsub(",", "", USDA$Sodium)
USDA$Potassium <-gsub(",", "", USDA$Potassium)
USDA$Sodium <- as.numeric(USDA$Sodium)
USDA$Potassium <- as.numeric(USDA$Potassium)
```

#### 4. Remove records (rows) with missing values in more than 4 attributes (columns). How many records remain in the data frame? (6 points)
```{r}
USDA <- USDA[(apply (is.na(USDA),1,sum)) <= 4,]
nrow(USDA)
#The remaining records are 6,887.
```

#### 5. For records with missing values for Sugar, Vitamin E and Vitamin D, replace missing values with mean value for the respective variable. (6 points)
```{r}
USDA$Sugar[is.na(USDA$Sugar)] = mean(USDA$Sugar[!is.na(USDA$Sugar)])

USDA$VitaminE[is.na(USDA$VitaminE)] = mean(USDA$VitaminE[!is.na(USDA$VitaminE)])


USDA$VitaminD[is.na(USDA$VitaminD)] = mean(USDA$VitaminD[!is.na(USDA$VitaminD)])

#checking if 0
#USDA$Sugar[is.na(USDA$Sugar)] 
#USDA$VitaminE[is.na(USDA$VitaminE)]
#USDA$VitaminD[is.na(USDA$VitaminD)]
```

#### 6. With a single line of code, remove all remaining records with missing values. Name the new Data Frame "USDAclean". How many records remain in the data frame? (6 points)
```{r}
USDAclean <- USDA[complete.cases(USDA), ]
nrow(USDAclean)
#The remaining records are 6,310.
```

#### 7. Which food has the highest sodium level? (6 points)
```{r}
max(USDAclean$Sodium)
USDAclean$Description[USDAclean$Sodium ==max(USDAclean$Sodium)]
#SALT,TABLE has the highest sodium level that is 38,758.
```

#### 8. Create a histogram of Vitamin C distribution in foods. (6 points)
```{r}
hist(USDAclean$VitaminC , xlim=c(0,100), breaks= 100, xlab = "Vitamin C", main ="Vitamin C Distribution")
```

#### 9. Create a boxplot to illustrate the distribution of values for TotalFat, Protein and Carbohydrate. (6 points)
```{r}
boxplot(USDAclean$TotalFat, USDAclean$Protein, USDAclean$Carbohydrate,main="Boxplot Illustration", names=c("TotalFat", "Protein","Carbs."))
```

#### 10. Create a scatterplot to illustrate the relationship between a food's TotalFat content and its Calorie content. (6 points)
```{r}
plot(USDAclean$TotalFat, USDAclean$Calories, main ="Scatterplot Illustration",xlab="TotalFat (green)",ylab="Calories (red)", col=c(2,3))
```

#### 11. Add a variable to the data frame that takes value 1 if the food has higher sodium than average, 0 otherwise. Call this variable HighSodium. Do the same for High Calories, High Protein, High Sugar, and High Fat. How many foods have both high sodium and high fat? (8 points)
```{r}
#High Sodium
USDAclean$HighSodium[USDAclean$Sodium > mean(USDAclean$Sodium)] <- 1
USDAclean$HighSodium[USDAclean$Sodium <= mean(USDAclean$Sodium)] <- 0

#High Calories
USDAclean$HighCalories[USDAclean$Calories > mean(USDAclean$Calories)] <- 1
USDAclean$HighCalories[USDAclean$Calories <= mean(USDAclean$Calories)] <- 0

#High Protein
USDAclean$HighProtein[USDAclean$Protein > mean(USDAclean$Protein)] <- 1
USDAclean$HighProtein[USDAclean$Protein <= mean(USDAclean$Protein)] <- 0

#High Sugar
USDAclean$HighSugar[USDAclean$Sugar > mean(USDAclean$Sugar)] <- 1
USDAclean$HighSugar[USDAclean$Sugar <= mean(USDAclean$Sugar)] <- 0

#High Fat
USDAclean$HighFat[USDAclean$TotalFat > mean(USDAclean$TotalFat)] <- 1
USDAclean$HighFat[USDAclean$TotalFat <= mean(USDAclean$TotalFat)] <- 0

High<-apply(USDAclean[c("HighSodium", "HighFat")], 1, sum)
table(High)
#644 foods have both high sodium and high fat.

```

#### 12. Calculate the average amount of iron, for high and low protein foods. (8 points)
```{r}
tapply(USDAclean$Iron, USDAclean$HighProtein, mean)
#The average amount of iron for high protein food is 3.069541.
#The average amount of iron for low protein food is 2.696634.
```

#### 13. Create a script for a "HealthCheck" program to detect unhealthy foods. Use the algorithm flowchart below as a basis for this script. (8 points)
```{r}
require(jpeg)
img<-readJPEG("HealthCheck.jpg")
plot(1:4, ty = 'n', ann = F, xaxt = 'n', yaxt = 'n')
rasterImage(img,1,1,4,4)
```
```{r}
HealthCheck <- function(food){if (food$HighSodium ==0) return ("Pass") else if (food$HighSugar ==0) return ("Pass") else if(food$HighFat ==0) return("Pass") else return("Fail")}
```

#### 14. Add a new variable called HealthCheck to the data frame using the output of the function. (8 points)
```{r}
for (index in 1:nrow(USDAclean)) {USDAclean$HealthCheck[index] = HealthCheck(USDAclean[index,])}
```

#### 15. How many foods in the USDAclean data frame fail the HealthCheck? (8 points)
```{r}
table(USDAclean$HealthCheck)
#237 foods fail the HealthCheck.
```

#### 16. Save your final data frame as "USDAclean_ [your last name]." (3 points)
```{r}
write.csv(USDAclean, "USDAclean_Park")
```


This is the end of Assignment 1

Ceni Babaoglu, PhD
