---
title: 'CMTH 642 Data Analytics: Advanced Methods'
output:
  html_document: default
  pdf_document: default
  word_document: default
---
<center> <h1> Assignment 3 (10%) </h1> </center>
<center>  <h3> [Ilhak Park] </h2> </center>
<center> <h3> [DJ0 & 501072432] </h2> </center>
---


#### 1. Import to R the following fiel: http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv (The dataset is related to white Portuguese "Vinho Verde" wine. For more info: https://archive.ics.uci.edu/ml/datasets/Wine+Quality) (3 points)
```{r}
wine <- read.csv(file="http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv", header = TRUE, sep =";")
```

#### 2. Check the datatypes of the attributes. (3 points)
```{r}
sapply (wine, class)
```
  
#### 3. Are there any missing values in the dataset? (4 points)
```{r}
str(wine)
sum(is.na(wine))
#there is no missing values in the dataset
```

#### 4. What is the correlation between the attributes other than Quality? (10 points)
```{r}
corWine <- cor(wine[-12])
corWine
library(corrplot)
corrplot(corWine, method="circle")
```

#### 5. Graph the frequency distribution of wine quality by using Quality. (10 points)
```{r}
hist(wine$quality)
```

#### 6.	Reduce the levels of rating for quality to three levels as high, medium and low. Assign the levels of 3 and 4 to level 0; 5 and 6 to level 1; and 7,8 and 9 to level 2. (10 points)
```{r}
for (i in  1:4898){ if (wine$quality[i]==3 | wine$quality[i]==4) {wine$quality[i] = 'Level 0: Low'} else if (wine$quality[i]==5 | wine$quality[i]==6) {wine$quality[i] = 'Level 1: Medium'} else if (wine$quality[i]==7 | wine$quality[i]==8 | wine$quality[i]==9) {wine$quality[i] = 'Level 2: High'}}

table(wine$quality)
```

#### 7. Normalize the data set by using the following function: (12 points)
```{r}
normalize <- function(x){
  return ((x - min(x)) / (max(x) - min(x)))
}
```

```{r}
wine_n <- as.data.frame(lapply(wine[-12], normalize))
wine_n <- cbind(wine_n, wine$quality)
summary(wine_n)
```

#### 8. Divide the dataset to training and test sets. (12 points)
```{r}
set.seed(1)
index <- sample (1:nrow(wine_n), 0.65 *nrow(wine_n))
wine_train <- wine_n[index,]
wine_test <- wine_n[-index,]
wine_train_labels <- wine_train[,12]
wine_test_labels <- wine_test[,12]
table(wine_train_labels)
table(wine_test_labels)

```

#### 9. Use the KNN algorithm to predict the quality of wine using its attributes. (12 points)
```{r}
library(class)
wine_test_pred <- knn(train = wine_train[,1:11], test = wine_test[,1:11], cl = wine_train[,12], k=10)
table(wine_test_pred)
```

#### 10. Display the confusion matrix to evaluate the model performance. (12 points)
```{r}
CM <- table(Actual = wine_test_labels, Predicted = wine_test_pred)
CM
```

#### 11. Evaluate the model performance by computing Accuracy, Sensitivity and Specificity. (12 points)
```{r}
library(e1071)
library(caret)
confusionMatrix(CM)
```

This is the end of Assignment 3

Ceni Babaoglu, PhD